---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Cure Model Pipeline

<!-- badges: start -->
<!-- badges: end -->

The goal of this package is to provide an easy way to build, evaluate and visualize mixture cure models ([link to paper draft](https://docs.google.com/document/d/1Lhom9H7Hxb8_C6d55Trs9JHSfJCn1i6yuu_xsDyYqZQ/edit?usp=sharing))


## Available Functions

- `fit_cure()` - wrapper for `smcure::smcure()` function that allows you to pass multilevel variables as factors directly (instead of first creating dummy codes). It also saves outputs necessary to create nomogram.
- `nomogram()-` - accepts an object created by `fit_cure()` and creates graphical nomogram based on model fit
- `cure_calibration()` - accepts an object created by `fit_cure()` and creates a calibration curve for uncured patients
- `multiple_mod_runs()`- allows you to check stability of models but running them multiple times and viewing distributions of variables and p-values. 
- Coming soon:
  - k-index and C-index wrapper functions
  - tidyr function to better clean/view coefficients from cure model

## Example 

```{r, eval = FALSE}
library(smcure)
library(ISwR)
library(gtsummary)
library(tidyverse)

# Prepare Data --------------------------
mel <- ISwR::melanom %>%
  mutate(status = case_when(status %in% c(1, 3) ~ 1, 
                            TRUE ~ 0)) %>%
  mutate(thick_cat = 
              case_when(thick <= 129 ~ "≤ 129", 
                        thick <= 322 ~ "less equal 322", 
                        thick > 322 ~ "greater 322")) %>%
  mutate(thick_cat = fct_relevel(thick_cat, "≤ 129", "≤ 322")) %>%
  mutate(sex = case_when(sex == 2 ~ "male", 
                         sex == 1 ~ "female"),
         ulc = case_when(ulc == 1 ~ "present", 
                         ulc == 2 ~ "absent"))

# Run Pipeline --------------------------
fit <- fit_cure(formula = Surv(days, status) ~ ulc + sex + thick_cat + thick ,
                data = mel)
cure_nomogram(fit, prediction_time = 300)
cure_calibration(fit, prediction_time = 300)


# Check Stability ----
formula <- Surv(days, status) ~ ulc + sex
x <- multiple_mod_runs(formula,
                              nboot = 200,
                              eps = 0.0001,
                              num_repeats = 3,
                              data = mel)
x$model_results

x$var_surv_stab
x$p_surv_stab

x$var_cure_stab
x$p_cure_stab

```


