---
title: "Cure Model Analysis Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cure Model Analysis Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
gt_compact_fun <- function(x) {
  gt::tab_options(x, 
                  table.font.size = 'small',
                  data_row.padding = gt::px(1),
                  summary_row.padding = gt::px(1),
                  grand_summary_row.padding = gt::px(1),
                  footnotes.padding = gt::px(1),
                  source_notes.padding = gt::px(1),
                  row_group.padding = gt::px(1))
}

library(smcure)
library(survival)
library(tidyverse)
library(gtsummary)
library(survival)
library(ggpubr)
library(survminer)
library(gt)

files <- list.files(here::here("R"), full.names = TRUE)

map(files, ~source(.x))
```


# Introduction

This package allows you to easily build and report cure mixture models. The modeling functions are wrappers for the `smcure()` function from the {smcure} package and several functions are available to diagnose and visualize your cure model results.

# Example data set

We will be using publicly available data on 205 stage I melanoma patients at the Department of Plastic Surgery, University Hospital of Odense, Denmark from 1962 to 1977. The data is available in the {ISwR} package (Dalgaard, 2008). 

Variables include:
- `ulc` - ulceration status
- `thick` - tumor thickness in mm
- `sex` - sex

We created categorized version of tumor thickness

```{r}
df <- ISwR::melanom

mel <- ISwR::melanom %>%
  mutate(status = case_when(status %in% c(1) ~ 1, 
                            status %in% c(2, 3) ~ 0)) %>%
  mutate(thick_cat = 
              case_when(thick <= 129 ~ "≤ 129", 
                        thick <= 322 ~ "less equal 322", 
                        thick > 322 ~ "greater 322")) %>%
  mutate(thick_cat = fct_relevel(thick_cat, "≤ 129", "≤ 322")) %>%
  mutate(sex = case_when(sex == 2 ~ "male", 
                         sex == 1 ~ "female"),
         ulc = case_when(ulc == 1 ~ "present", 
                         ulc == 2 ~ "absent"))

mel %>%
  select(thick_cat, sex, ulc) %>%
  tbl_summary() %>%
  bold_labels()


```

To calculate overall survival, we will only consider death from melanoma as an event. Patients who died from other cause or were lost to follow-up are censored. 

```{r}
fit <- survfit(Surv(days/365, status) ~ ulc, data = mel)

fit %>% ggsurvplot()

```

# Fitting the Cure Model

The `fit_cure()` function is a wrapper for `smcure::smcure()`. It differs from the original function in that it allows you to pass multi-level categorical variables and also provides necessary model outputs to create calibration curves and calculate k-indicies. 


```{r }

fit <- fit_cure(formula = Surv(days, status) ~ ulc + sex + 
                  thick_cat,
                data = mel)

x <- mmcure(fit$smcure_model_object) %>%
  transmute(variable = str_remove_all(name, fixed("Z[, -1]")), 
            value, 
            pvalue) %>%
  filter(variable != "(Intercept)")


y <- mmsurv(fit$smcure_model_object) %>%
  transmute(variable = str_remove_all(
    str_remove_all(name, fixed("X[, -1]")), fixed("`")), 
            value, 
            pvalue) 

all <- x %>% left_join(y, by = "variable", suffix = c("_cure", "_survival"))
all %>%
  gt()
```

# K-Index

Below we calculate the K and C indicies for this model

```{r}
library(ROCR)
  
mod_data <- mel %>%
  mutate(sex = as.integer(as.factor(sex))-1) 

mm <- model.matrix(~ulc + sex + 
                  thick_cat, data = mod_data)[, -1]

mod2 <- mod_data %>%
  dplyr::select(days, status) %>%
  cbind(mm)
  
k_fit <- evacure::smcure1(Surv(days, status) ~ ulcpresent + 
                            sex + 
                            `thick_catgreater 322` + 
                            `thick_catless equal 322`,
      cureform = ~ ulcpresent + 
                            sex + 
                            `thick_catgreater 322` + 
                            `thick_catless equal 322`,
      data = mod2,
      model = "ph",
      nboot = 200,
      Var = T
      )
```

```{r}

k_index <- k_fit$eva["K"]
c_index <- k_fit$eva["C"]


tribble(
  ~"Statistic", ~"Value", 
  "k-index", k_index, 
  "c-index", c_index) %>%
  mutate_if(is.numeric, ~round(.x, 3)) %>%
  gt() %>%
  gt::tab_options(table.font.size = "small", 
        data_row.padding = gt::px(1),
        summary_row.padding = gt::px(1), 
        grand_summary_row.padding = gt::px(1),
        footnotes.padding = gt::px(1), 
        source_notes.padding = gt::px(1),
        row_group.padding = gt::px(1))
```


# Calibration Curve

```{r}

cure_calibration(fit, prediction_time = 1000)
```

