---
title: "Cure Model Analysis Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cure Model Analysis Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
gt_compact_fun <- function(x) {
  gt::tab_options(x, 
                  table.font.size = 'small',
                  data_row.padding = gt::px(1),
                  summary_row.padding = gt::px(1),
                  grand_summary_row.padding = gt::px(1),
                  footnotes.padding = gt::px(1),
                  source_notes.padding = gt::px(1),
                  row_group.padding = gt::px(1))
}

library(smcure)
library(survival)
library(dplyr)
library(gtsummary)
library(survival)
library(survminer)
library(gt)
library(purrr)
library(cureit)
library(ggsurvfit)
library(forcats)

```


# Introduction

This package allows you to easily build and report cure mixture models. The modeling functions are wrappers for the `smcure()` function from the {smcure} package and several functions are available to diagnose and visualize your cure model results.

# Example data set

We will be using publicly available data on 205 stage I melanoma patients at the Department of Plastic Surgery, University Hospital of Odense, Denmark from 1962 to 1977. The original data is available in the {ISwR} package (Dalgaard, 2008). A version of this data in available in the package `cureit::melanoma`.

Variables include:

- `ulceration`: ulceration status
- `tumor_thickness`: tumor thickness in mm
- `sex`: sex

A summary of the main available variables is below:

```{r}
melanoma %>%
  select(tumor_thickness, sex, ulceration) %>%
  tbl_summary() %>%
  bold_labels()
```

To calculate overall survival, we will only consider death from melanoma as an event. Patients who died from other cause or were lost to follow-up are censored. 

Disease death outcome data is plotted and summarized below using the {ggsurvfit} package and {gtsummary}:

```{r}
fit <- survfit(Surv(years, status) ~ 1, data = melanoma)

fit %>% 
  ggsurvfit::ggsurvfit() + 
  ylab("Disease-related Death Probability")+
  xlab("Years") + 
  ylim(0, 1) +
  theme_classic()

```

From the Kaplan-meier curves we can see the disease death rate plateaus around .7, indicating that not all patients may experience disease-related death. The patients that do experience this event, may be different from those who do not in a way that would be useful to categorize.

Cure models can help tease apart the contribution of clinical variables to these potentially heterogeneous sub-cohorts. 

# Fitting the Cure Model

The `cureit()` function is a wrapper for `smcure::smcure()` that allows you to code cure models using a tidy framework. Below, we will fit a model using ulceration status, sex and tumor thickness as variables of interest in both cure and survival portions. 

Note: You can optionally provide different sets of covariates to the survival portion and the cure portion:

```{r}
fit <- cureit(surv_formula = Surv(days, status) ~ ulceration + sex + tumor_thickness,
              cure_formula = ~ ulceration + sex + tumor_thickness, data = melanoma)
```

## Producing Table Summary of the Cure Model

```{r}
tidy_mods <- tidy(fit)

tbl_regression(fit, tidy_fun = tidy) %>% 
  bold_labels()

# cure model tidy table
tidy_cure_df <- fit$tidy$df_cure

# survival model tidy table
tidy_surv_df <- fit$tidy$df_surv
```

## Model Prediction

```{r}
p <- cureit(Surv(days, status) ~ ulceration + sex + tumor_thickness,
  cure_formula = ~ ulceration + sex + tumor_thickness,
  data = melanoma)

x <- predict(p, times = seq(100, 5000, 100), newdata = melanoma, brier = TRUE)

# names of lists that can be extracted from x
names(x)
```

# Nomogram

```{r}
nomogram(fit, time = 3)
```

