---
title: "Cure Model Analysis Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cure Model Analysis Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = TRUE
)
gt_compact_fun <- function(x) {
  gt::tab_options(x, 
                  table.font.size = 'small',
                  data_row.padding = gt::px(1),
                  summary_row.padding = gt::px(1),
                  grand_summary_row.padding = gt::px(1),
                  footnotes.padding = gt::px(1),
                  source_notes.padding = gt::px(1),
                  row_group.padding = gt::px(1))
}

library(smcure)
library(survival)
library(dplyr)
library(gtsummary)
library(survival)
library(survminer)
library(gt)
library(purrr)
library(cureit)
library(ggsurvfit)
library(forcats)

```


# Introduction

{cureit} allows you to easily build and report mixture cure models. Functions are available to fit, diagnose, and visualize your cure model results. More information on theory and application of cure models using {cureit} in the supporting manuscript: 

> Whiting K, Fei T, Singer S, Qin, LX. Cureit: An End-to-End Pipeline for Implementing Mixture Cure Models With an Application to Liposarcoma Data. JCO Clin Cancer Inform 8, e2300234(2024).[DOI:10.1200/CCI.23.00234](https://ascopubs.org/doi/abs/10.1200/CCI.23.00234)

# Example data set

We will be using publicly available data on 205 stage I melanoma patients at the Department of Plastic Surgery, University Hospital of Odense, Denmark from 1962 to 1977. The original data is available in the {ISwR} package (`ISwR::melanom`, Dalgaard, 2008). A version of this data is available in the cureit package `cureit::melanoma`.

Variables include:

- `ulceration`: ulceration status
- `tumor_thickness`: tumor thickness in mm
- `sex`: sex

The main available variables in this data set are summarized below using the {gtsummary} package:

```{r}
melanoma %>%
  select(tumor_thickness, sex, ulceration) %>%
  tbl_summary() %>%
  bold_labels()
```

To calculate overall survival, we will only consider death from melanoma as an event. Patients who died from other cause or were lost to follow-up are censored. 

Disease death outcome data is plotted and summarized below using the {ggsurvfit} package and {gtsummary}:

```{r}
fit <- survfit(Surv(years, status) ~ 1, data = melanoma)

fit %>% 
  ggsurvfit::ggsurvfit() + 
  ylab("Disease-related Death Probability")+
  xlab("Years") + 
  ylim(0, 1) +
  theme_classic()

```

From the Kaplan-meier curves we can see the disease death rate plateaus around .7, indicating that not all patients may experience disease-related death. The patients that do experience this event, may have different clinical charcteristics from those who do not. Cure models can help characterize the contribution of clinical variables to these potentially heterogeneous sub-cohorts. 

# Fitting the Cure Model

The `cureit()` function is a wrapper for `smcure::smcure()` and allows you to fit cure models and estimate coefficients and confidence intervals (via bootstrapping). Results can be summarized and organized using a tidy framework (see {broom} package for more information). Below, we will fit a cure model using ulceration status, sex and tumor thickness as variables of interest in both cure and survival portions. 

Note: You can optionally provide different sets of covariates to the survival portion and the cure portion:

```{r}
fit <- cureit(
  surv_formula = Surv(days, status) ~ ulceration + sex + tumor_thickness,
  cure_formula = ~ ulceration + sex + tumor_thickness, data = melanoma
)
```

## Producing Table Summary of the Cure Model

A tidy method is available to turn model estimates into a data frame of model info.


```{r}
tidy(fit)
```
Additionally, you can extract tidy fits of both sub models separately as well: 

```{r }

# cure model tidy table
fit$tidy$df_cure

# survival model tidy table
fit$tidy$df_surv
```
A basic `gtsummary::tbl_summary()` method for `cureit` model objects is also available **(please note: this is experimental and will be updated soon to include improved formatting of submodel components):**

```{r}

tbl_regression(fit, exponentiate = TRUE) %>% 
  bold_labels() %>%
  bold_p()
```

## Model Prediction

Make predictions on cureit model: 

```{r}
p <- cureit(Surv(days, status) ~ ulceration + sex + tumor_thickness,
  cure_formula = ~ ulceration + sex + tumor_thickness,
  data = melanoma)

x <- predict(p, times = seq(100, 5000, 100), newdata = melanoma, brier = TRUE)

# names of lists that can be extracted from x
names(x)
```

# Nomogram

Create a nomogram based on the model fit: 

```{r}
nomogram(fit, time = 3)
```

