---
title: "Analyze Variant Data"
output: html_document
date: "2023-02-15"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
library(cureit)
library(ggsurvfit)
library(labelled)
library(gtsummary)
library(survival)
library(tidycmprsk)
library(tidyverse)
theme_gtsummary_compact()
load(here::here("inst", "deid_data_2022-09-02[1526].RData"))
df <- df_sel
# Summary ------
fit <- survfit(Surv(os_months, dod) ~ 1, data = df_sel)

m <- median(df$age_presentation)

df <- df %>%
  mutate(age_cat = case_when(
    age_presentation <= m ~ glue::glue("age <= {m}"), 
    age_presentation > m ~ glue::glue("age > {m}"), 
    TRUE ~ NA_character_
  ))
```

```{r }
ggsurvfit(fit) + ylim(0, 1) + 
  add_risktable() + theme_classic()
```

```{r}
process_reg_table <- function(tbl_reg) {
    
  tbl_reg$table_body <- tbl_reg$table_body %>%
    separate_wider_delim(
      label,
      names = c("label", "label2"), 
             delim = ",")%>%
    separate_wider_delim(
      variable,
      names = c("variable", "variable2"), 
             delim = ",")
  
  cure <- tbl_reg
  cure$table_body <- cure$table_body %>%
    filter(label2 == " Cure model") %>%
    filter(label != "(Intercept)")
  
  surv <- tbl_reg
  surv$table_body <- surv$table_body%>%
    filter(label2 == " Survival model")
  
  tbl_merge(list(cure, surv), 
            tab_spanner = c("Cure Model",
                            "Survival Model"))
}
```


This report uses data from Bartlett et al. 2019 to model death from disease, local recurrence, and distant recurrence using mixture cure models. 

The original paper modeled these three endpoints using competing risks methods to analyze prognostic significance of histologic subtypes in liposarcoma.  

Since the cohort has a high proportion of individuals who do not experience the event, mixture cure models may be more suited to explicitly model survival as a mixture of two types of patients: those who are "cured" (will not experience event during follow-up time) and those who are not. Using mixture cure models, we can estimate the proportion of long term survivors and also describe the time to event among those who are not "cured". Using this formulation, we can investigate patient characteristics that are related to each of the model components. 

# Cohort Summaries


```{r}
df <- df %>%
  mutate(site2 = case_when(
    site1_name %in% c("Lower extremity", "Upper extremity") ~ "Lower/upper Extremity", 
    TRUE ~ site1_name))
df <- df %>%
  set_variable_labels(.labels = kwiktools::names_to_labels(.))
var_label(df) <- list(
  sex_dscrp = "Sex",
  depth_dscrp = "Depth",
  size_category_dscrp = "Size Categorized"
)
df_cure <- df %>%
  mutate(dod = case_when(death == 1 ~ 1, 
                         (death == 2 | death == 0) ~ 0, 
                         TRUE ~ NA_real_)) 
#%>%
#  filter(resection_category != "Unknown")

df <- df_cure
```

```{r}
df %>%
  select(
         variant_name_specify,
         age_presentation, 
         sex_dscrp,
         site1_name, 
         max_tumor_size,
         size_category_dscrp, 
         depth_dscrp, 
         resection_category, 
         bone_invasion, 
         chemo_sequence, 
         rt_sequence) %>%
  tbl_summary(by = "variant_name_specify") %>% 
  bold_labels() %>%
  add_p() %>% 
  add_overall() %>%
  bold_p()
```

# Endpoints 


```{r}
df %>%
  select(death) %>% 
  tbl_summary(label = death ~ "Death (1 is disease-related)") %>%
  bold_labels()
```

<br>

```{r}
df %>%
  select(lr_y_n) %>% 
  tbl_summary(label = lr_y_n ~ "Local Recurrence") %>%
  bold_labels()
```

<br> 

```{r}
df %>%
  select(dr_y_n) %>% 
  tbl_summary(label = dr_y_n ~ "Distant Recurrence") %>%
  bold_labels()
```

<br>



# Univariable Cure Model


```{r}

uni_func <- function(var, df_input, nboot) {
  df <- df_input %>% 
    select("os_months",
           "dod",
           var) %>%
  na.omit()
  

  f_surv <- as.formula(paste0("Surv(os_months, dod) ~ ", var))
  f_cure <- as.formula(paste0("~ ", var))
  
  
  p <- cureit(surv_formula = f_surv,
            cure_formula = f_cure,
            data = df,
            nboot = nboot)
  p

}
```

```{r, eval=FALSE}
nboot = 500

age_uni <- uni_func("age_presentation", df,
                    nboot = 500)
save(age_uni, file = here::here("inst",
                         "cure-model-fits",
                         "age_uni_2023"))
age_uni$tidy

age_cat_uni <- uni_func("age_cat", df,
                    nboot = 100)
save(age_cat_uni, file = here::here("inst",
                         "cure-model-fits",
                         "age_uni_2023"))
age_cat_uni$tidy




v_uni <- uni_func("variant_name_specify", df,
                  nboot = 500)
save(v_uni, file = here::here("inst",
                         "cure-model-fits",
                         "v_uni_2023"))
v_uni$tidy

sex_uni <- uni_func("sex_dscrp", df,
                    nboot = 500)
save(sex_uni, file = here::here("inst",
                         "cure-model-fits",
                         "sex_uni_2023"))

site_uni <- uni_func("site1_name", df,
                    nboot = 500)
save(site_uni, file = here::here("inst",
                         "cure-model-fits",
                         "site_uni_2023"))

maxsize_uni <- uni_func("max_tumor_size", df,
                    nboot = 500)
save(maxsize_uni, file = here::here("inst",
                         "cure-model-fits",
                         "maxsize_uni_2023"))

dep_uni <- uni_func("depth_dscrp", df,
                    nboot = 500)
save(dep_uni, file = here::here("inst",
                         "cure-model-fits",
                         "dep_uni_uni_2023"))

load(here::here("inst",
                         "cure-model-fits",
                         "dep_uni_uni_2023"))

df2 <- df %>%
  filter(resection_category != "Unknown")
res_uni <- uni_func("resection_category", df2,
                    nboot = 500)
save(res_uni, file = here::here("inst",
                         "cure-model-fits",
                         "res_uni_2023"))


df2 <- df %>%
  mutate(log_tumor_size = log(max_tumor_size, 2))

logsize_uni <- uni_func("log_tumor_size", df2,
                    nboot = 100)

save(logsize_uni, file = here::here("inst",
                         "cure-model-fits",
                         "logsize_uni_2023"))

```

```{r}

model_fits <- list.files(
  here::here("inst", "cure-model-fits"), full.names = TRUE)

for(i in 1:length(model_fits)) load(model_fits[[i]]) 
```

```{r }
reg_tbl <- map(list(age_uni, v_uni, sex_uni,
                    site_uni, maxsize_uni, logsize_uni, res_uni),
               ~tbl_regression(.x, 
                               exponentiate = TRUE, 
                               tidy_fun = cureit::tidy))

x <- tbl_stack(reg_tbl) %>%
  bold_labels() %>%
  bold_p()

x$table_body <- x$table_body %>%
  separate_wider_delim(
    label,
    names = c("label", "label2"), 
           delim = ",")%>%
  separate_wider_delim(
    variable,
    names = c("variable", "variable2"), 
           delim = ",")

cure <- x

cure$table_body <- cure$table_body%>%
  filter(label2 == " Cure model") %>%
  filter(label != "(Intercept)")

cure

surv <- x

surv$table_body <- surv$table_body%>%
  filter(label2 == " Survival model")

surv

tbl_merge(list(cure, surv), 
          tab_spanner = c("Cure Model",
                          "Survival Model"))



```


<br>

# Multivariable Models

## Subtype, Size


```{r, eval=FALSE}
multi1  <- cureit(surv_formula = Surv(os_months, dod) ~ variant_name_specify + 
                  log(max_tumor_size, 2),
                  cure_formula = ~ variant_name_specify + 
                  log(max_tumor_size, 2),
                data = df_cure,
                nboot = 50)
#save(multi1, file =  here::here("inst", "cure-model-fits", "multi1.Rd"))

multi20  <- cureit(surv_formula = Surv(os_months, dod) ~ variant_name_specify + 
                  max_tumor_size,
                  cure_formula = ~ variant_name_specify + 
                  max_tumor_size,
                data = df_cure,
                nboot = 10)

multi50 %>% 
  tbl_regression(tidy_fun = tidy, exponentiate = TRUE) %>%
  process_reg_table()

save(multi20, file =  here::here("inst", "cure-model-fits", "multi20.Rd"))
```

```{r }
load(here::here("inst", "cure-model-fits", "multi1.Rd"))

z <- multi1 %>% 
  tbl_regression(tidy_fun = cureit::tidy,
                 exponentiate = TRUE)

process_reg_table(z)
```

```{r}

x$table_body <- x$table_body %>%
  separate_wider_delim(
    label,
    names = c("label", "label2"), 
           delim = ",")%>%
  separate_wider_delim(
    variable,
    names = c("variable", "variable2"), 
           delim = ",")

cure <- x

cure$table_body <- cure$table_body%>%
  filter(label2 == " Cure model") %>%
  filter(label != "(Intercept)")

cure

surv <- x

surv$table_body <- surv$table_body%>%
  filter(label2 == " Survival model")

surv

tbl_merge(list(cure, surv), 
          tab_spanner = c("Cure Model",
                          "Survival Model"))

```

### Brier

```{r}

null_index <- purrr::map_lgl(multi1$smcure$bootstrap_fit, is.null)
error_count <- sum(null_index)
multi1$smcure$bootstrap_fit <- multi1$smcure$bootstrap_fit[!null_index]
    
pred <- predict(object = multi1,
                method = "prob",
                times = seq(0,380,10),
                brier = TRUE,
                cox = TRUE)

pred 
plot(seq(10,390,10), pred$brier,type="s",xlab="Time",ylab="Brier score",ylim=c(0,0.15))
lines(seq(10,390,10), pred$brier_cox,type="s",col="red")
legend("topleft", c("Cure model","Cox model"),col=c("black","red"),lty=1)

pred1 <- pred

```

```{r}
times <- seq(1,340,10)
bootbrier2 <- Brier_inference_bootstrap(object = multi1, times = times)
save(bootbrier2, file = here::here("inst", "cure-model-fits", "brierboot.Rd"))
times <- times

plot.new()
plot(times,bootbrier2$brier,type="s",xlab="Time",ylab="Brier score")
lines(times,bootbrier2$brier_2.5,type="s",lty=2)
lines(times,bootbrier2$brier_97.5,type="s",lty=2)
lines(times,bootbrier2$brier_cox,type="s",col="red")
lines(times,bootbrier2$brier_cox_2.5,type="s",lty=2,col="red")
lines(times,bootbrier2$brier_cox_97.5,type="s",lty=2,col="red")
legend("topleft",c("Cure model","Cox model"),col=c("black","red"),lty=1)
```


```{r}

bb <- bootbrier2[c(1:8)] %>%
  bind_cols() %>%
  mutate(times = times)


bb_s <- bb %>%
  select(times, brier, brier_cox) %>%
  pivot_longer(-times)

bb_ci_min <- bb %>%
  select(times, "brier" = brier_2.5, "brier_cox" = brier_cox_2.5) %>%
  pivot_longer(-times)

bb_ci_max <- bb %>%
  select(times, "brier" = brier_97.5, "brier_cox" = brier_cox_97.5) %>%
  pivot_longer(-times)


bb_l <- bb_s %>%
  left_join(bb_ci_min, by = c("times", "name"), suffix = c("_score", "_ci_min")) %>%
  left_join(bb_ci_max, by =  c("times", "name")) %>%
  rename("value_ci_max" = value) %>%
  mutate(Model =
           case_when(
             name == "brier" ~ "Brier: Cure Model", 
             name == "brier_cox" ~ "Brier: Cox Model"))
 
bb_l %>%
  ggplot() + 
  geom_step(aes(x = times/12, y = value_score, color = Model)) +
  geom_ribbon(aes(x = times/12, ymin = value_ci_min,
                  ymax = value_ci_max, fill = Model), alpha = .3) +
  theme_classic() + 
  ylab("Brier Score") +
  xlab("Time (Years)") + 
  xlim(c(0, 300/12)) + ylim(0, .15) 



```

## Subtype, Size, Depth


```{r}
multi2  <- cureit(surv_formula = Surv(os_months, dod) ~ variant_name_specify + 
                    depth_dscrp +
                  log(max_tumor_size, 2),
                  cure_formula = ~ variant_name_specify + depth_dscrp +
                  log(max_tumor_size, 2),
                data = df_cure,
                nboot = 500)

multi2 %>% tbl_regression(tidy_fun = tidy, exponentiate = TRUE)

save(multi2, file =  here::here("inst", "cure-model-fits", "multi2.Rd"))
```


## Subtype, Size, Depth, Site


```{r}
multi3  <- cureit(surv_formula = Surv(os_months, dod) ~
                    variant_name_specify + 
                    depth_dscrp +
                  log(max_tumor_size, 2) + 
                    site1_name,
                  
                  cure_formula = ~ variant_name_specify +
                    depth_dscrp +
                  log(max_tumor_size, 2) +
                    site1_name,
                data = df_cure,
                nboot = 500)

multi3 %>% tbl_regression(tidy_fun = tidy, exponentiate = TRUE) %>%
  process_reg_table()

save(multi3, file =  here::here("inst", "cure-model-fits", "multi3.Rd"))
```



## Subtype, Size, Age Cat


```{r}
multi4  <- cureit(surv_formula = Surv(os_months, dod) ~
                    variant_name_specify + 
                  log(max_tumor_size, 2) + 
                    age_cat,
                  
                  cure_formula = ~ variant_name_specify +
                    depth_dscrp +
                  log(max_tumor_size, 2) +
                    age_cat,
                data = df_cure,
                nboot = 100)

multi4 %>% tbl_regression(tidy_fun = tidy, exponentiate = TRUE) %>%
  process_reg_table()

save(multi3, file =  here::here("inst", "cure-model-fits", "multi3.Rd"))
```

## Nomogram
```{r}
# # This causes error
# df_up <- df_cure %>%
#   mutate(`Log Tumor Size` =  log(max_tumor_size, 2), 
#          `Subtype` = variant_name_specify)
#   
# multi1_nomo  <- cureit(surv_formula = Surv(os_months, dod) ~ `Subtype` + 
#                   `Log Tumor Size`,
#                   cure_formula = ~ `Subtype` + 
#                   `Log Tumor Size`,
#                 data = df_up,
#                 nboot = 0)
# 
# cureit::nomogram(x = multi1_nomo, time = 12*5)

df_up <- df_cure %>%
  mutate(log_tumor_size =  log(max_tumor_size, 2),
         subtype = variant_name_specify)

multi1_nomo  <- cureit(surv_formula = Surv(os_months, dod) ~ subtype +
                 log_tumor_size,
                  cure_formula = ~ subtype +
                  log_tumor_size,
                data = df_up,
                nboot = 0)

cureit::nomogram(x = multi1_nomo, time = 12*5, angle = 45, size = .1)



```


# Local Recurrence

```{r}
df_lr <- df_cure %>%
  filter(!(resection_category %in% c("R2"))) %>%
  filter(is.na(rt_sequence))

omit <- df_cure %>%
  filter(resection_category %in% c("R2") | !is.na(rt_sequence))

df_lr <- df_lr %>%
  mutate(lr_y_n = case_when(lr_y_n == "Yes" ~ 1 , 
                           lr_y_n == "No" ~ 0))
```

```{r}
df_lr %>% 
  select(death, lr_y_n) %>% 
  tbl_cross()
```


<br>

```{r}
fit <- survfit2(Surv(lrfs_months/12, lr_y_n) ~ variant_name_specify, 
                data = df_lr)

ggsurvfit(fit) + ylim(0, 1) + 
  add_risktable(risktable_stats = "n.risk") +
  theme_classic() + 
  xlab("Years") + 
  ylab("Local Recurrence-free\n Survival") +
  add_confidence_interval() + 
  add_pvalue(location = "annotation") +
  theme(legend.position = "top")
```

```{r}

lr_uni_func <- function(var, df_input, nboot) {
  df <- df_input %>% 
    select("lrfs_months",
           "lr_y_n",
           var) %>%
  na.omit()
  

  f_surv <- as.formula(paste0("Surv(lrfs_months, lr_y_n) ~ ", var))
  f_cure <- as.formula(paste0("~ ", var))
  
  
  p <- cureit(surv_formula = f_surv,
            cure_formula = f_cure,
            data = df,
            nboot = nboot)
  p

}
```

```{r}
lr_v_uni <- lr_uni_func("variant_name_specify", df_lr,
                  nboot = 500)

save(lr_v_uni, file = here::here("inst",
                         "cure-model-fits",
                         "lr_v_uni_2023"))

```

```{r}
# 
# model_fits <- list.files(
#   here::here("inst", "cure-model-fits"), full.names = TRUE)
# 
# for(i in 1:length(model_fits)) load(model_fits[[i]]) 
```

```{r }
reg_tbl <- map(list(lr_v_uni),
               ~tbl_regression(.x, 
                               exponentiate = TRUE, 
                               tidy_fun = cureit::tidy))

x <- tbl_stack(reg_tbl) %>%
  bold_labels() %>%
  bold_p()

x$table_body <- x$table_body %>%
  separate_wider_delim(
    label,
    names = c("label", "label2"), 
           delim = ",")%>%
  separate_wider_delim(
    variable,
    names = c("variable", "variable2"), 
           delim = ",")

cure <- x

cure$table_body <- cure$table_body%>%
  filter(label2 == " Cure model") %>%
  filter(label != "(Intercept)")

cure

surv <- x

surv$table_body <- surv$table_body%>%
  filter(label2 == " Survival model")

surv

tbl_merge(list(cure, surv), 
          tab_spanner = c("Cure Model",
                          "Survival Model"))



```


